### 2009-09-21 latex/paper1/contents.tex
versao revisada.

% TODO: say that we use DIRECTED GRAPHS

@@

\section{Abstract}

Software clustering algorithms automatically recognize a system's modular
structure by analyzing its implementation.  Due to the lack of well document
software systems, though, the issue of testing these algorithms is still
underexplored, limiting both their adoption in the industry and the development
of better algorithms.  We propose to rely on software models to produce
arbitrarily large test sets. In this paper we consider three such models and
analyze how similar the artifacts they produce are from artifacts from real
software systems.

@@

Therefore, to optimize the development, a development manager can decompose the
system into modules, or groups of components, so that there are as few
inter-module dependencies as possible.% (i.e., the modules are weakly couple.).
Each module is, then, threated as an unit of work assignment \cite{Parnas1972}:
each module is assigned to one team and, because of the little inter-module
dependency, the communication between teams is reduced, therefore speeding up
the development.

Therefore, to decompose a system, a manager needs to have global knowledge of
the dependencies between its components. Although dependency analysis tools can
extract such information from the source code of the system, the task of finding
modules can be overwhelming because of the large number of components and
dependencies.

Software clustering algorithms, also known as architecture recovery algorithms
\cite{Mancoridis1998,Anquetil1999,Tzerpos2000,Andritsos2005},
%\cite{Pollet2007}, 
automatically decompose a software system modules by analyzing the network of
dependencies between its components.  Although many algorithms have been
proposed, there are few comparative studies.
%empirical studies regarding the quality of the decompositions they find.
%EVALUATION CRITERION
One important test for a software clustering algorithm is to apply it to a
collection of software systems with known reference decompositions made by
experienced developers \cite{Anquetil1999}. It is expected that a good algorithm
finds decompositions that are similar to reference decompositions, which can be
measured by a similarity metric such as MoJo \cite{Tzerpos1999} or EdgeSim
\cite{Mitchell2001}.

Unfortunately there are few publicly available reference decompositions
\cite{Koschke2000} and, as a result, few comparative studies exist. Also,
because it is costly to obtain reference decompositions for large systems, most
studies test the algorithms with a couple of small and medium systems
\cite{Anquetil1999,Maqbool2007,Bittencourt2009}.

@@

Of course, it is desirable that the synthetic networks resemble networks
extracted from real software systems. In this paper, we present three network
models and show empirically that all three models are capable of synthesizing
networks that resemble software networks. We consider object-oriented systems
with classes as components.

The paper is organized as follows:
Section \ref{sec:models} presents three network models.
%
Section \ref{sec:characterization} demonstrates a method to tell if a network
resembles a software network.
%
Section \ref{sec:evaluation} describes an experiment that shows that the three
models synthesize networks that resemble software networks.
%
Section \ref{sec:conclusion} discusses the impact of this research and future
work.

@@

Network theory studies general properties of many types of networks by using
statistical analysis. In the last decade, it has been found that many networks
arising from sociology, biology, technology and other domains present remarkable
structural similarities. It has been shown that, in these networks, the
distribution of vertex degrees is a power law, i.e., the number of vertices
connected to $k$ edges is proportional to $k^{-\gamma}$, where $\gamma$ is a
positive constant. These networks are called scale-free networks
\cite{Barabasi1999}.

Network theory has been applied to class dependency networks and it was shown
that they are also scale-free networks \cite{Myers2003,Valverde2002}. Thus, the
study of the structure of object-oriented software systems can benefit from the
vast literature available about scale-free networks.

In particular, many scale-free network models were proposed. These models are
algorithms that were proven, either formally or empirically, to synthesize
networks that are scale-free. In this section we present three models that
synthesize scale-free networks with built-in modular decompositions: BCR+, CGW,
and LF. Their output can be used to test software clustering algorithms.

@@

The BCR model \cite{Bollobas2003} aims to model the network of hyperlinks
between web pages as a directed graph without modules. We have developed an
extension to this model, called BCR+, that adds modules to the construction of
the network. The model accepts the following parameters:

@@

The parameter $\mu$ controls the proportion external edges in the network. Lower
values of $\mu$ lead to networks with fewer external edges and, therefore, to
weakly coupled modules.

@@

Because the model is described in detail in the original paper, only an
intuitive notion of the meaning of the parameters is provided here. This model
is similar to BCR+ in which it grows an initial network by sucessive
applications of operations. In this case, there four operations, which are
chosen according to probabilities $p_1$, $p_2$, $p_3$, and $p_4$:

@@

The constant $\alpha$ controls the proportion of edges that connect vertices
from distinct modules. For $\alpha = -1$, most edges will connect distinct
modules. For $\alpha > 0$, most edges will connect vertices in the same module,
and the greater the $\alpha$, more strong is the tendency. For $\alpha = 0$
there is no bias, the two kinds of edges are equaly likely.

@@

There is also a special version of the model in which the edges weights are
discarded and the modules are non overlapping. We used the original
implementation of this version, available at
\url{http://santo.fortunato.googlepages.com/inthepress2}. It accepts the
following parameters:

@@

Following the work by Milo et al. \cite{Milo2004}, similarity between two
networks can be measured by computing Pearson's correlation coefficient between
the corresponding TCPs, which yields a value between -1 (most dissimilar) and 1
(most similar):

@@

\subsubsection{Non-software networks} We have collected 66 networks from distinct
domains, such as biology, sociology, technology, and linguistics. These networks
are freely available on the Internet and have been used in previous researches.
CITE

% S-score

@@

$$
\mathrm{S\mbox{-}score}(a) = \frac{
\displaystyle\sum_{s \in S} \mathrm{similarity}(a, s)
}{|S|} \mathrm{,}
$$

@@

Two networks, though, showed high S-score, above 0.83. The network of links
between blogs on politics showed S-score 0.97. The neural network of the worm C.
Elegans also showed a high S-score (0.88). Further investigation is needed in
order to discover the reasons behind the high values and whether auxiliary
metrics can differentiate these networks from software networks.

@@

Increasing the threshold has the effect of reducing the recall, because fewer
software networks are classified as software-like. Decreasing the threshold has
the effect of reducing the precision, since more non-software networks are
classified as software-like.

@@

In the previous section it was shown that many networks, although scale-free,
can be distinguished from software networks by a simple classification model
based on triad concentration profiles. In this section we show empirically that
the three network models previously presented synthesize networks that are
indistinguishable from software networks.
The experiment consists of synthesizing networks using many combinations of
parameters from the three models, and then classifying each network as
software-like or non software-like. %After that, we present rules for the choice
%of parameter values that lead to software-like networks.

\subsection{Synthetic Data Set}

We want to investigate if, with a proper choice of parameters, a model is
capable of synthesizing a network that resembles software networks. Because the
possible combinations of parameter values are infinite, we have set the number
of vertices to 1000 and then varied the remaining parameters in discrete steps.
In this section we describe the combinations of parameters values used for each
model.

@@

We have chosen five different module graphs, which where extracted from actual
dependencies between Java archives distributed with five different software
systems of our sample: GEF (2 archives), iBATIS (4 archives), MegaMek (8
archives), findbugs (16 archives), and zk (32 archives). Because many of these
archives are intended to be reused in distinct projects, they provide a good
coarse-grained approximation of the concept of module.

@@

The algorithm 1R from data mining was used to help discover such relationship.
It analyzes the parameters and the classification of each network and finds a
rule that relates the value of one single parameter with the classification. We
are interested in rules for networks classified as software-like and in the
accuracy of these rules, i.e., the proportion of networks that are correctly
classified. The rules found by 1R are shown in Table \ref{tab:rules}.

@@

The rules are very simple and, thus, easy to follow. Despite their simplicity,
they have high accuracy, approximately 80\% in all models.  
%These rules allow one to know if a particular combination of parameter values
%is likely to lead to software-like networks.

@@

In a future work, we intend to use synthetic networks in the evaluation of
software clustering algorithms that were previously studied with real networks
\cite{Wu2005}. After that we will be able to compare the results obtained by
the approaches.

>>>>

%   ok: abstract
%   ok: say that we use DIRECTED GRAPHS
% TODO: review

@@

@@

Therefore, to optimize the development of a system, a manager can decompose the
system into modules, or groups of components, so that there are as few
inter-module dependencies as possible. % (i.e., the modules are weakly couple.).
%Each module is, then, threated as an unit of work assignment \cite{Parnas1972}:
Then, each module is assigned to one team and, because of the little inter-module
dependency, the communication between teams is reduced, therefore speeding up
the development.

To decompose a system into modules, a manager needs to have global knowledge of
the dependencies between its components. Although dependency analysis tools can
extract such information from the source code of the system, the task of finding
a good decomposition can be overwhelming because of the large number of
components and dependencies that need to be considered.

Software clustering algorithms, also known as architecture recovery algorithms
\cite{Mancoridis1998,Anquetil1999,Tzerpos2000,Andritsos2005},
%\cite{Pollet2007}, 
automatically decompose a software system by analyzing the network of
dependencies between its components. %Although many algorithms have been
%proposed, few comparative studies exist.  empirical studies regarding the
%quality of the decompositions they find.  EVALUATION CRITERION
One important test for a software clustering algorithm is to apply it to a
collection of software systems with known reference decompositions made by
experienced developers \cite{Anquetil1999}. It is expected that a good algorithm
finds decompositions that are similar to reference decompositions, which can be
measured by a similarity metric such as MoJo \cite{Tzerpos1999} or EdgeSim
\cite{Mitchell2001}.

Unfortunately, there are few publicly available reference decompositions
\cite{Koschke2000} and, as a result, there are few studies that compare software
clustering algorithms. Also, because it is costly to obtain reference
decompositions for large systems, most studies test the algorithms with a couple
of small and medium systems \cite{Anquetil1999,Maqbool2007,Bittencourt2009}.

@@

A desirable property of synthetic networks is that they resemble their natural
counterparts. In this paper, we present three network models and show
empirically that all three models are capable of synthesizing networks that
resemble software networks. We consider object-oriented systems with classes as
components.

The paper is organized as follows: Section \ref{sec:models} presents three
network models.
%
Section \ref{sec:characterization} demonstrates a method to tell if a network
resembles a software network.
%
Section \ref{sec:evaluation} shows empirically that the three models synthesize
networks that resemble software networks.
%
Section \ref{sec:conclusion} discusses the impact of this research and future
work.

@@

Network theory studies general properties of many types of networks by using
statistical analysis. In the last decade, remarkable similarities have been
found in networks arising from domains as diverse as sociology, biology,
technology, and linguistics. It has been shown that, in these networks, the
distribution of vertex degrees is a power law, i.e., the number of vertices
connected to $k$ edges is proportional to $k^{-\gamma}$, where $\gamma$ is a
positive constant. These networks are called scale-free networks
\cite{Barabasi1999}.

Network theory has been applied to class dependency networks, represented as
directed unweighted graphs, and it was shown that they are also scale-free
networks \cite{Myers2003,Valverde2002}. Thus, the study of the structure of
object-oriented software systems can benefit from the vast literature available
about scale-free networks.

In particular, many scale-free network models were proposed. These models are
algorithms that were proven, either formally or empirically, to synthesize
networks that are scale-free. In this section we present three models that
synthesize scale-free networks with built-in modular decompositions: BCR+, CGW,
and LF. The networks are represented as directed unweighted graphs with modules,
which makes them suitable for testing software clustering algorithms.

@@

The BCR model \cite{Bollobas2003} synthesizes directed scale-free networks
without modules. We have developed an extension to this model, called BCR+, that
adds modules to the construction of the network. The BCR+ model accepts the
following parameters:

@@

The parameter $\mu$ controls the proportion of external edges in the network.
Lower values of $\mu$ lead to networks with fewer external edges. 
%and, therefore, to weakly coupled modules.

@@

Because the model is described in detail in the original paper, only an
intuitive notion of the meaning of the parameters is provided here. This model
is similar to BCR+ in which it grows an initial network by sucessive
applications of operations. In this case, there are four operations, which are
chosen according to probabilities $p_1$, $p_2$, $p_3$, and $p_4$:

@@

The constant $\alpha$ controls the proportion of edges that connect vertices
from distinct modules. For $\alpha = -1$, most edges will connect distinct
modules. For $\alpha > 0$, most edges will connect vertices in the same module,
and the greater the $\alpha$, the stronger the tendency. For $\alpha = 0$ there
is no bias, the two kinds of edges are equaly likely.

@@

There is also a special version of the model in which the edges weights are
discarded and the modules do not overlap. We used the original implementation of
this version, available at
\url{http://santo.fortunato.googlepages.com/inthepress2}. It accepts the
following parameters:

@@

Following another work by Milo et al. \cite{Milo2004}, similarity between two
networks can be measured by computing Pearson's correlation coefficient between
the corresponding TCPs, which yields a value between -1 (most dissimilar) and 1
(most similar):

@@

\subsubsection{Non-software networks} We have collected 66 networks from
distinct domains, such as biology, sociology, technology, and linguistics, with
size ranging from 32 to 18163 vertices. These networks are freely available on
the Internet and have been previously studied in the literature.
% S-score

@@

$$
\mathrm{S\mbox{-}score}(a) ~=~ \frac{
\displaystyle\sum_{s \in S} \mathrm{similarity}(a, s)
}{|S|} \mathrm{,}
$$

@@

Two networks, though, showed high S-score: the network of links between blogs on
politics, with S-score 0.97, and the neural network of the worm C. Elegans, with
S-score 0.88. Further investigation is needed in order to discover the reasons
behind the high values and whether auxiliary metrics can differentiate these
networks from software networks.

@@

Increasing the threshold has the effect of reducing the recall, because fewer
software networks are classified as software-like. Decreasing the threshold has
the effect of reducing the precision, because more non-software networks are
classified as software-like.

@@

In the previous section it was shown that many networks, although scale-free,
can be distinguished from software networks by a simple classification model
based on triad concentration profiles. In this section we show empirically that
the three network models previously presented can synthesize networks that are
indistinguishable from software networks. The experiment consists of
synthesizing networks using many combinations of parameters from the three
models, and then classifying each network as software-like or non software-like.
%After that, we present rules for the choice of parameter values that lead to
%software-like networks.  \subsection{Synthetic Data Set}
%
%We want to investigate if, with a proper choice of parameters, a model is
%capable of synthesizing a network that resembles software networks. 
Because the possible combinations of parameter values are infinite, we have set
the number of vertices to 1000 and then varied the remaining parameters in
discrete steps.
%In this section we describe the combinations of parameters values used for each
%model.

@@

We have chosen five different module graphs, which where extracted from actual
dependencies between Java archives distributed with five different software
systems of our sample: GEF (2 archives), iBATIS (4 archives), MegaMek (8
archives), findbugs (16 archives), and zk (32 archives). Because many of these
archives are intended to be reused in distinct projects, they provide a good
coarse-grained approximation to the concept of module.

@@

The algorithm 1R \cite{OneR} from machine learning was used to help discover
such relationship. It analyzes the parameters and the classification of each
network and finds a rule that relates the value of one single parameter with the
classification. Such rules can be evaluated according to their accuracy, i.e.,
the proportion of networks that are correctly classified. The rules found by 1R
are shown in Table \ref{tab:rules}.

@@

The rules are very simple and, thus, easy to follow. Despite their simplicity,
they have high accuracy, approximately 80\% for all models.  
%These rules allow one to know if a particular combination of parameter values
%is likely to lead to software-like networks.

@@

In a future work, we intend to use synthetic networks in the evaluation of
software clustering algorithms that were previously tested with real networks.
After that we will be able to compare the results obtained by different
approaches.
